\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-03-11 paper.qmd
Rewrite model summary section

The prediction approach involves employing the K-nearest neighbors (KNN) algorithm. In this method, which operates as an unsupervised classification problem, the nearest high-D model point (averaged high-D point) is identified for the new high-D data point. Once the nearest high-D model point is identified, its coordinates are used as the predicted 2D embedding for the new high-D data point. This step is particularly valuable due to the limitations of certain NLDR techniques, like tSNE, which don't provide a straightforward method for predicting 2D embedding. As a result, our approach presents a solution that can generate predicted 2D embedding, irrespective of the NLDR technique employed, effectively addressing this gap in functionality.

The concept of "residuals" is pivotal in evaluating the accuracy of representing bin centroids in high dimensions. To quantify this accuracy, we introduce an error metric, which measures the sum of squared differences between the high-dimensional data ($x_{ij}$) and the predicted bin centroid data in high-dimensional space ($C_{x_ij}$) across all bins and dimensions. Mathematically, this error is expressed as:

@@

Here, $n$ represents the number of observations, $p$ represents the dimensions, $x_{ij}$ is the actual high-dimensional data, and $C_{x_ij}$ is the predicted bin centroid data in high dimensions.

The error metric outlined above provides valuable insights into the overall accuracy of our predictive model. By quantifying the squared deviations between the actual and predicted values across all bins and dimensions, we gain a comprehensive understanding of how well our method captures and represents the underlying structure of the data in the reduced 2D space. This assessment is crucial for evaluating the efficacy of our NLDR technique in preserving the essential information present in the original high-dimensional data.

@@

Moving on to the assessment of prediction accuracy, we calculate the Mean Squared Error (MSE). The MSE helps measure the average squared differences between the actual high-dimensional data ($x_{ij}$) and the predicted bin centroid data in high-D ($C_{x_ij}$) values across all bins. Mathematically, this is expressed as:

@@

Here, $b'$ signifies the total number of bins without empty bins, $p$ denotes the number of dimensions in the high-dimensional data, and $n$ represents the number of observations.

Additionally, we gauge the model's performance using the Akaike Information Criterion (AIC), calculated by the formula:

@@

These metrics, MSE and AIC, collectively offer valuable insights into the model's predictive performance, considering both accuracy and complexity in the predictions.

@@

>>>>

The prediction approach involves performing the K-nearest neighbors (KNN) algorithm for an unsupervised classification problem. First, the nearest high-D model point is identified for a given new high-D point. Then, the corresponding 2D centroid mapping for the identified high-D model point is determined. Finally, the coordinates of this 2D centroid are used as the predicted 2D embedding for the new high-D data point. This step is particularly valuable due to the limitations of some NLDR techniques, like tSNE, which don't provide a straightforward method for prediction. As a result, our approach offers a solution that capable of generating predicted 2D embedding regardless of the NLDR technique employed, effectively addressing this functional gap.

Residuals are essential for evaluating the accuracy of representing high-D points by the high-D mapping of 2D bin centroids. To measure this accuracy, an error metric is introduced, quantifying the sum of squared differences between the high-D data ($x_{ij}$) and the high-D mapping of the 2D bin centroid data ($C_{x_ij}$) across all observations and dimensions (see @eq-equation11).

@@

Here, $n$ represents the number of observations, $p$ represents the dimensions of high-D data, $x_{ij}$ is the high-D data, and $C_{x_ij}$ is the high-D mapping of the 2D bin centroid.

<!--
The concept of "residuals" is pivotal in evaluating the accuracy of representing bin centroids in high-D. To quantify this accuracy, we introduce an error metric, which measures the sum of squared differences between the high-D data ($x_{ij}$) and the predicted bin centroid data in high-D space ($C_{x_ij}$) across all bins and dimensions. Mathematically, this error is expressed as:

Here, $n$ represents the number of observations, $p$ represents the dimensions, $x_{ij}$ is the high-D data, and $C_{x_ij}$ is the predicted bin centroid data in high-D.

The error metric outlined above provides valuable insights into the overall accuracy of our predictive model. By quantifying the squared deviations between the actual and predicted values across all bins and dimensions, we gain a comprehensive understanding of how well our method captures and represents the underlying structure of the data in the reduced 2D space. This assessment is crucial for evaluating the efficacy of our NLDR technique in preserving the essential information present in the original high-dimensional data.
-->

@@

To assess how well our method captures and represents the underlying structure of the high-D data, Mean Squared Error (MSE) and Akaike Information Criterion (AIC) are used. When computing MSE, total model error (see @sec-summary) is divided by the number of observations to make it as a mean value (see @eq-equation9).

@@

Furthermore, to incorporate with the parameters in our model, which is the non-empty bins, and as a statistical measure used for model selection and comparison, AIC is used (see @eq-equation10).

@@

Here, $b'$ signifies the number of bins non-empty bins, $p$ denotes the number of dimensions in the high-D data, and $n$ represents the number of observations.

@@

In conclusion, our proposed model offers a novel approach for visualizing high-dimensional (high-D) data by leveraging non-linear dimension reduction (NLDR) techniques. Through a series of steps including hexagonating NLDR data, triangulating bin centroids, and lifting the 2D triangular mesh into high dimensions, our model effectively transforms complex high-D data into interpretable 2D representations. The model's performance is evaluated using goodness of fit statistics such as Mean Squared Error (MSE) and Akaike Information Criterion (AIC), providing insights into its accuracy and reliability. Overall, our model presents a valuable tool for researchers and practitioners in various fields to gain deeper insights from high-dimensional datasets.

Our algorithm mainly consists...

These Goodness of fit statistics useful in encountering the prediction accuracy when we predict the 2D embedding for the new high-D data. Because, we try to find the nearest high-D mappings as the first step of prediction which will participate the prediction process. When we need to find which different NLDR technique and which parameter choices is giving the best representation of high-D data, MSE and AIC are also useful.

### 2024-03-11 paper.tex
Rewrite model summary section

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, breakable, boxrule=0pt, enhanced, borderline west={3pt}{0pt}{shadecolor}, interior hidden, sharp corners]}{\end{tcolorbox}}\fi

@@

The prediction approach involves employing the K-nearest neighbors (KNN)
algorithm. In this method, which operates as an unsupervised
classification problem, the nearest high-D model point (averaged high-D
point) is identified for the new high-D data point. Once the nearest
high-D model point is identified, its coordinates are used as the
predicted 2D embedding for the new high-D data point. This step is
particularly valuable due to the limitations of certain NLDR techniques,
like tSNE, which don't provide a straightforward method for predicting
2D embedding. As a result, our approach presents a solution that can
generate predicted 2D embedding, irrespective of the NLDR technique
employed, effectively addressing this gap in functionality.

The concept of ``residuals'' is pivotal in evaluating the accuracy of
representing bin centroids in high dimensions. To quantify this
accuracy, we introduce an error metric, which measures the sum of
squared differences between the high-dimensional data (\(x_{ij}\)) and
the predicted bin centroid data in high-dimensional space (\(C_{x_ij}\))
across all bins and dimensions. Mathematically, this error is expressed
as:

@@

Here, \(n\) represents the number of observations, \(p\) represents the
dimensions, \(x_{ij}\) is the actual high-dimensional data, and
\(C_{x_ij}\) is the predicted bin centroid data in high dimensions.

The error metric outlined above provides valuable insights into the
overall accuracy of our predictive model. By quantifying the squared
deviations between the actual and predicted values across all bins and
dimensions, we gain a comprehensive understanding of how well our method
captures and represents the underlying structure of the data in the
reduced 2D space. This assessment is crucial for evaluating the efficacy
of our NLDR technique in preserving the essential information present in
the original high-dimensional data.

@@

Moving on to the assessment of prediction accuracy, we calculate the
Mean Squared Error (MSE). The MSE helps measure the average squared
differences between the actual high-dimensional data (\(x_{ij}\)) and
the predicted bin centroid data in high-D (\(C_{x_ij}\)) values across
all bins. Mathematically, this is expressed as:

@@

Here, \(b'\) signifies the total number of bins without empty bins,
\(p\) denotes the number of dimensions in the high-dimensional data, and
\(n\) represents the number of observations.

Additionally, we gauge the model's performance using the Akaike
Information Criterion (AIC), calculated by the formula:

@@

These metrics, MSE and AIC, collectively offer valuable insights into
the model's predictive performance, considering both accuracy and
complexity in the predictions.

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, sharp corners, boxrule=0pt, interior hidden, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable]}{\end{tcolorbox}}\fi

@@

The prediction approach involves performing the K-nearest neighbors
(KNN) algorithm for an unsupervised classification problem. First, the
nearest high-D model point is identified for a given new high-D point.
Then, the corresponding 2D centroid mapping for the identified high-D
model point is determined. Finally, the coordinates of this 2D centroid
are used as the predicted 2D embedding for the new high-D data point.
This step is particularly valuable due to the limitations of some NLDR
techniques, like tSNE, which don't provide a straightforward method for
prediction. As a result, our approach offers a solution that capable of
generating predicted 2D embedding regardless of the NLDR technique
employed, effectively addressing this functional gap.

Residuals are essential for evaluating the accuracy of representing
high-D points by the high-D mapping of 2D bin centroids. To measure this
accuracy, an error metric is introduced, quantifying the sum of squared
differences between the high-D data (\(x_{ij}\)) and the high-D mapping
of the 2D bin centroid data (\(C_{x_ij}\)) across all observations and
dimensions (see Equation~\ref{eq-equation11}).

@@

Here, \(n\) represents the number of observations, \(p\) represents the
dimensions of high-D data, \(x_{ij}\) is the high-D data, and
\(C_{x_ij}\) is the high-D mapping of the 2D bin centroid.

@@

To assess how well our method captures and represents the underlying
structure of the high-D data, Mean Squared Error (MSE) and Akaike
Information Criterion (AIC) are used. When computing MSE, total model
error (see Section~\ref{sec-summary}) is divided by the number of
observations to make it as a mean value (see
Equation~\ref{eq-equation9}).

@@

Furthermore, to incorporate with the parameters in our model, which is
the non-empty bins, and as a statistical measure used for model
selection and comparison, AIC is used (see
Equation~\ref{eq-equation10}).

@@

Here, \(b'\) signifies the number of bins non-empty bins, \(p\) denotes
the number of dimensions in the high-D data, and \(n\) represents the
number of observations.

@@

In conclusion, our proposed model offers a novel approach for
visualizing high-dimensional (high-D) data by leveraging non-linear
dimension reduction (NLDR) techniques. Through a series of steps
including hexagonating NLDR data, triangulating bin centroids, and
lifting the 2D triangular mesh into high dimensions, our model
effectively transforms complex high-D data into interpretable 2D
representations. The model's performance is evaluated using goodness of
fit statistics such as Mean Squared Error (MSE) and Akaike Information
Criterion (AIC), providing insights into its accuracy and reliability.
Overall, our model presents a valuable tool for researchers and
practitioners in various fields to gain deeper insights from
high-dimensional datasets.

Our algorithm mainly consists\ldots{}

These Goodness of fit statistics useful in encountering the prediction
accuracy when we predict the 2D embedding for the new high-D data.
Because, we try to find the nearest high-D mappings as the first step of
prediction which will participate the prediction process. When we need
to find which different NLDR technique and which parameter choices is
giving the best representation of high-D data, MSE and AIC are also
useful.
